---
title: "Introduction to ggplot2"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

A review of penguin dataframe.
```{r}
penguins
```

Scatterplot that shows all penguins and them flipper length and body mass.
```{r}
ggplot(data = penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```

Add species as colored points.
```{r}
ggplot(data = penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point()
```

Add smooth regresion line to each specie.
```{r}
ggplot(data = penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  geom_smooth(method = "lm")
```

Pass the responsability of color points to point geom, so that, at *local level*.
```{r}
ggplot(data = penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(mapping = aes(color = species)) +
  geom_smooth(method = "lm")
```

For the legibility of the plot, i can add other shapes for each specie. This is 
useful for people with color blindness and other differences of color vision.
```{r}
ggplot(data = penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")
```

For change label texts i use the function `labs()`.
```{r}
ggplot(data = penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm") +
  labs(
    title = "Body mass and flipper length", 
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins", 
    x = "Flipper length (mm)", 
    y = "Body mass (g)", 
    color = "Species", 
    shape = "Species"
  )
```











