---
title: "Introduction to ggplot2"
output: html_notebook
---

# Libraries
```{r}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```

# Basic instructions
A review of penguin dataframe.
```{r}
penguins
```

Scatterplot that shows all penguins and them flipper length and body mass.
```{r}
ggplot(data = penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point()
```

Add species as colored points.
```{r}
ggplot(data = penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point()
```

Add smooth regresion line to each specie.
```{r}
ggplot(data = penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point() +
  geom_smooth(method = "lm")
```

Pass the responsability of color points to point geom, so that, at *local level*.
```{r}
ggplot(data = penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(mapping = aes(color = species)) +
  geom_smooth(method = "lm")
```

For the legibility of the plot, i can add other shapes for each specie. This is 
useful for people with color blindness and other differences of color vision.
```{r}
ggplot(data = penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm")
```

For change label texts i use the function `labs()`.
```{r}
ggplot(data = penguins, 
       mapping = aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(mapping = aes(color = species, shape = species)) +
  geom_smooth(method = "lm") +
  labs(
    title = "Body mass and flipper length", 
    subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins", 
    x = "Flipper length (mm)", 
    y = "Body mass (g)", 
    color = "Species", 
    shape = "Species"
  )
```


# Different plots

Explore the different `binwidth` is useful for discover better patterns in the 
data distribution.
```{r}
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 20)
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 2000)
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(binwidth = 200)
```

An exploration using *species*.
```{r}
ggplot(data = penguins, mapping = aes(y = species)) + 
  geom_bar()
```

If want change the color bars, there two variants: `color` change the edges 
of the bars, `fill` change the color of entire bar.
```{r}
ggplot(penguins, aes(x = species)) +
  geom_bar(color = "dodgerblue")

ggplot(penguins, aes(x = species)) +
  geom_bar(fill = "dodgerblue")
```

If want change the bins amount, i use the `bins` parameter.
```{r}
ggplot(penguins, mapping = aes(x = body_mass_g)) +
  geom_histogram(bins = 20, color = "black", fill = "lightgreen")
```

Exploring the **diamonds dataset** with histogram and different binwidths.
```{r}
one = .11
two = .12
three = .13
ggplot(diamonds, mapping = aes(x = carat)) + 
    geom_histogram(binwidth = one, color = "black", fill = "dodgerblue") + 
    labs(
      title = paste("Binwidth of ", one)
    )
ggplot(diamonds, mapping = aes(x = carat)) + 
    geom_histogram(binwidth = two, color = "black", fill = "dodgerblue") + 
    labs(
      title = paste("Binwidth of ", two)
    )
ggplot(diamonds, mapping = aes(x = carat)) + 
    geom_histogram(binwidth = three, color = "black", fill = "dodgerblue") + 
    labs(
      title = paste("Binwidth of ", three)
    )
```

Also use the density plot.
```{r}
ggplot(diamonds, mapping = aes(x = carat)) + 
    geom_density()
```









