---
title: "Plots to Visualize Relationships"
output: html_notebook
---

```{r}
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
```


# A numerical and a categorical variable

The **box plot** can help to visualize the general distribution of data using 
numerical variables related to categorical variables.
```{r}
ggplot(penguins, aes(x = species, y = body_mass_g)) +
  geom_boxplot() +
  labs(
    title = "Spread of body mass for each specie", 
    x = "Species", 
    y = "Body mass [g]"
  )
```

Other plot to visualize the relationship is **density plot**. The `alpha` 
parameter set the transparency of filled density curves. Also, the `linewidth` 
can set the thickness of line used for the curves.
```{r}
ggplot(penguins, aes(x = body_mass_g, color = species, fill = species)) +
  geom_density(alpha = 0.5, linewidth = 0.7) + 
  labs(
    title = "Density curves for each specie", 
    x = "Body mass [g]", 
    y = "Density"
  )
```

# Two categorical variables

The **bar plot** is used to visulize the relationship among two categorical 
variables. There are two setup:
+ For absolute frequencies.
+ For relative frequencies.

When i want see the **exact numbers of individuals** the absolute frequency bar plot 
is more desirable. The problem is it's distribution in each island not clearly.
```{r}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar() +
  labs(
    title = "Species in each island", 
    x = "Island", 
    y = "Absolute frequency"
  )
```

When i want see the proportion of individuals in each group the relative 
frequency bar plot is more desirable.
```{r}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(position = "fill") +
  labs(
    title = "Species in each island", 
    x = "Island", 
    y = "Relative frequency"
  )
```

# Two numerical variables

If the two variables are numerical the **scatterplot** used to visualize the 
data.
```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point() + 
  labs(
    title = "Two variables of penguins", 
    x = "Flipper length [mm]", 
    y = "Body mass [g]"
  )
```


# Three or more variables

When the number of variables is upper 2, the strategy focus in represent those 
variables with colors or shapes.
```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point(aes(color = species, shape = island)) + 
  labs(
    title = "Multiple characteristics of penguins", 
    x = "Flipper length [mm]", 
    y = "Body mass [g]"
  )
```

However, the graph begin to be seen crowded of information. For enhance this 
visualization, i use the `facet_wrap` that split the graph in subplots based on 
categorical variable.
```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + 
  geom_point(aes(color = species)) + 
  facet_wrap(~island) + 
  labs(
    title = "Subplots based on penguin island", 
    x = "Flipper length [mm]", 
    y = "Body mass [g]"
  )
```

# The applications of relationships

```{r}
glimpse(mpg)
```


Visualize the relationship among *highway miles per gallon* `hwy` and 
*engine displacement [L]* `displ`. This use *city miles per gallon* `cty` as the 
third variable.
```{r}
ggplot(data = mpg, mapping = aes(x = hwy, y = displ)) +
  geom_point(aes(color = cty, size = cty)) + 
  labs(
    title = "Distribution of mechanic characteristics", 
    x = "Highway miles per gallon", 
    y = "Engine displacement [L]", 
    color = "City miles per gallon", 
    size = "City miles per gallon"
  )
```

If use the same third variable to `linewidth` parameter, this cause errors in 
the graph.
```{r}
ggplot(data = mpg, mapping = aes(x = hwy, y = displ)) +
  geom_point()
```

Using the penguin data frame, inspecting the relationship among: 
`bill_depth_mm`, `bill_length_mm` and `species`.
```{r}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + 
  geom_point(aes(color = species)) + 
  labs(
    title = "Penguin bill attributes", 
    x = "Bill depth [mm]", 
    y = "Bill length [mm]", 
    color = "Species"
  )
```

If i use faceting, the result is:
```{r}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + 
  geom_point(aes(color = species)) + 
  facet_wrap(~species) + 
  labs(
    title = "Penguin bill attributes", 
    x = "Bill depth [mm]", 
    y = "Bill length [mm]", 
    color = "Species"
  )
```

